AppPAL for Android

### 2015-05-28 paper.tex
Updated formatting

% vim: set nowrap tw=0:
\documentclass[a4paper]{scrartcl}
\usepackage{defaultstyle}
\usepackage{localstyle}

@@

\title{AppPAL for Android\\
Capturing and Checking App Installation Policies}
\author{Joseph Hallett \and David Aspinall}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
  % \begin{itemize}
  %   \item \comment{Sentence 1: State the problem}
  Users must judge apps based only on the information presented to them by the store.
  %   \item \comment{Sentence 2: Why is it an interesting problem}
  Employers rely on employees using company devices appropriately to enforce security policies at work.
  Some users spend time to pick apps with care others don't.
  This may cause frustration when they realise what data the apps have access to.
  %   \item \comment{Sentence 3: What does my solution achieve}
  We present AppPAL: a policy language that can enforce app installation policies.
  Policies use statements from third parties and delegation relationships.
  %   \item \comment{Sentence 4: What follows from my solution}
  This gives us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
  % \end{itemize}
\end{abstract}

@@

App stores give some information about their apps; such as permissions and review scores.
Android apps show a confusing list of permissions (\autoref{fig:facebook}).
Users don't all understand how permissions relate to their device~\citep{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision which apps to use must be made by a human.

Not all apps are suitable. 
A large amount of \ac{pus} is being propagated for Android devices~\citep{Truong:2014bi,Svajcer:2013tp}. 
Employees are increasingly using their own phones for work (the \ac{byod} scheme.)
A employers may wish to restrict which apps their employees can use.
The \ac{ciso} or IT department may set a policy to prevent information leaks.
Some users worry apps will misuse their personal data;
  such a user avoids apps which can access their location, or address book.
They apply their own personal security policy when downloading apps.

@@

%\comment{My contribution is\ldots}
We present AppPAL, an authorization logic for reasoning about Android apps.
The language is an instantiation of SecPAL~\citep{Becker:2006vh} tailored to making statements about apps.
The language allows us to reason about apps using statements from third parties.
The implementation allows us to enforce the policies on a device.
We can express trust relationships amongst these parties; use constraints to perform additional checks.
This lets us enforce deeper and more complex policies than existing tools such as Kirin~\citep{Enck:2009ko}.

@@

Our research looks at the problem of picking software.
Given there are some apps you want and others you don't:
  how can you express your preferences in such a way that they can be enforced automatically?
How can we translate policy documents from natural language into something machine checkable?
Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?

@@

\begin{figure}
\begin{lstlisting}[numbers=left,
                   escapeinside={@}{@}]
"alice" says "emma" can-say inf App isInstallable. @\label{lst:corporate_1}@
"emma" says App isInstallable @\label{lst:corporate_2}@
  if "no-tracking-policy" isMetBy(App),
     "reputable-source-policy" isMetBy(App),
     "anti-virus-policy" isMetBy(App).
"emma" says "reputable-source-policy" isMetBy(App) @\label{lst:corporate_3}@
  if App isReputable.
"emma" says "google-play" can-say 0 App isReputable. @\label{lst:corporate_4}@
"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = False.
"emma" says "no-location-permissions" can-act-as "no-tracking-policy". @\label{lst:corporate_6}@
"emma" says "no-location-permissions" isMetBy(App) @\label{lst:corporate_7}@
  if App isAnApp
  where
    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
    hasPermission(App, "ACCESS_COURSE_LOCATION") = False.
\end{lstlisting}
\caption{AppPAL policy implementing Emma's security requirements}
\label{lst:corporate}
\end{figure}

@@

\begin{lstlisting}
"bob" says "no-locations-permissions" can-act-as "no-tracking-policy".
"bob" says "no locations-permissions" isMetBy(App)
  if App isAnApp
  where
    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
    hasPermission(App, "ACCESS_COARSE_LOCATION") = False.

"bob" says "location-taint-analysis" can-act-as "no-tracking-policy".
"bob" says "location-taint-analysis" isMetBy(App)
  if App isAnApp
  where
    taintDroidCheckLeak(App, "Location", "Internet") = False.
\end{lstlisting}

@@

In SecPAL there is no notion of type.
All entities are either constants (effectively strings) or variables.
There is no differentiation made between a \emph{voiced} constant (one who speaks),
and a \emph{subject} (one that is the subject of a verb-phrase, or predicate argument).
In AppPAL the subjects are usually apps or policies.
Apps do not utter facts.
Consequently there is no need to check them as potential speakers of \emph{can-say} statements.
Similarly if a constant never appears as a subject of a statement, as most speakers don't,
  there is no need to check it as a possible subject for a \emph{can-act-as} statement.
We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.

@@

Work by \citeauthor{Sadeh:2014vq} looked at modelling users' willingness to grant different kinds of permissions~\citep{Sadeh:2014vq}.
They found four clusters of different users each with different kinds of policies:
  conservatives who don't like granting many permissions,
  advanced users who seem to have more complex policies,
  unconcerned users who'll comfortably grant anything,
  and fence-sitters (about half of all users) who seemed not to really care the majority of the time.
This suggests that users are applying policies and do care how apps behave.
Further more \citeauthor{Sadeh:2014vq} speculate that the fence-sitters seemed not to care because of \emph{warning fatigue}.
This suggests a system like AppPAL that can reduce user interaction with permission warnings might be interesting.
Using a policy unconcerned and fence-sitting users could take up more privacy preserving policies without having to change their behavior.

>>>>

% vim: set nowrap tw=0:
\documentclass[twoside,letterpaper]{soups}

\pdfpagewidth=8.5truein{}
\pdfpageheight=11truein{}

\usepackage{graphicx}
\renewcommand{\topfraction}{0.99} % be more aggressive about text around floats
\renewcommand{\floatpagefraction}{0.99}
\pagestyle{plain} % page numbers

\usepackage{defaultstyle}
\usepackage{localstyle}

@@

\title{AppPAL for Android}
\subtitle{Capturing and Checking App Installation Policies}
\numberofauthors{2}
\author{\alignauthor
  Joseph Hallett \\ \email{J.Hallett@sms.ed.ac.uk}
  \alignauthor
  David Aspinall \\ \email{David.Aspinal@ed.ac.uk}}

\begin{document}

\maketitle{}

\begin{abstract}
  % \begin{itemize}
  %   \item \comment{Sentence 1: State the problem}
  Users must judge apps based only on the information presented to them by the store.
  %   \item \comment{Sentence 2: Why is it an interesting problem}
  Employers rely on employees using company devices appropriately to enforce security policies at work.
  Some users spend time to pick apps with care others do not.
  This may cause frustration when they realise what data the apps have access to.
  %   \item \comment{Sentence 3: What does my solution achieve}
  We present AppPAL: a policy language that can enforce app installation policies.
  Policies use statements from third parties and delegation relationships.
  %   \item \comment{Sentence 4: What follows from my solution}
  This gives us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
  % \end{itemize}
\end{abstract}

@@

App stores give some information about their apps; such as permissions and review scores.
Android apps show a confusing list of permissions (\autoref{fig:facebook}).
Users do not all understand how permissions relate to their device~\citep{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision which apps to use must be made by a human.

Not all apps are suitable.
A large amount of \ac{pus} is being propagated for Android devices~\citep{Truong:2014bi,Svajcer:2013tp}.
Employees are increasingly using their own phones for work (the \ac{byod} scheme.)
An employer may wish to restrict which apps their employees can use.
The \ac{ciso} or IT department may set a policy to prevent information leaks.
Some users worry apps will misuse their personal data;
  such a user avoids apps which can access their location, or address book.
They apply their own personal security policy when downloading apps.

@@

%\comment{My contribution is\ldots}
We present AppPAL, an authorization logic for reasoning about Android apps.
The language is an instantiation of SecPAL~\citep{Becker:2006vh} tailored to making statements about apps.
The language allows us to reason about apps using statements from third parties.
The implementation allows us to enforce the policies on a device.
We can express trust relationships amongst these parties; use constraints to do additional checks.
This lets us enforce deeper and more complex policies than existing tools such as Kirin~\citep{Enck:2009ko}.

@@

Our research looks at the problem of picking software.
Given there are some apps you want and others you do not:
  how can you express your preferences in such a way that they can be enforced automatically?
How can we translate policy documents from natural language into something machine checkable?
Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?

@@

\begin{figure}
\begin{lstlisting}[numbers=left,
                   escapeinside={@}{@}]
"alice" says "emma" can-say inf
  App isInstallable. @\label{lst:corporate_1}@

"emma" says App isInstallable @\label{lst:corporate_2}@
  if "no-tracking-policy" isMetBy(App),
     "reputable-source-policy" isMetBy(App),
     "anti-virus-policy" isMetBy(App).

"emma" says "reputable-source-policy" isMetBy(App) @\label{lst:corporate_3}@
  if App isReputable.

"emma" says "google-play" can-say 0 App isReputable. @\label{lst:corporate_4}@

"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = False.

"emma" says "no-location-permissions"
  can-act-as "no-tracking-policy". @\label{lst:corporate_6}@

"emma" says "no-location-permissions" isMetBy(App) @\label{lst:corporate_7}@
  if App isAnApp
  where
    hasPermission(App, "ACCESS_FINE_LOCATION") = false,
    hasPermission(App, "ACCESS_COURSE_LOCATION") = false.
\end{lstlisting}
\caption{AppPAL policy implementing Emma's security requirements}
\label{lst:corporate}
\end{figure}

@@

\begin{lstlisting}
"bob" says "no-locations-permissions"
  can-act-as "no-tracking-policy".

"bob" says "no-locations-permissions" isMetBy(App)
  if App isAnApp
  where
    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
    hasPermission(App, "ACCESS_COARSE_LOCATION") = False.

"bob" says "location-taint-analysis"
  can-act-as "no-tracking-policy".

"bob" says "location-taint-analysis" isMetBy(App)
  if App isAnApp
  where
    taintDroidCheckLeak(App, "Location", "Internet")
      = False.
\end{lstlisting}

@@

In SecPAL there is no notion of type.
All entities are either constants (effectively strings) or variables.
There is no differentiation made between a \emph{voiced} constant (one who speaks),
and a \emph{subject} (one that is the subject of a verb-phrase, or predicate argument).
In AppPAL the subjects are usually apps or policies.
Apps do not utter facts.
Consequently there is no need to check them as potential speakers of \emph{can-say} statements.
Similarly if a constant never appears as a subject of a statement, as most speakers do not,
  there is no need to check it as a possible subject for a \emph{can-act-as} statement.
We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.

@@

Work by \citeauthor{Sadeh:2014vq} looked at modelling users' willingness to grant different kinds of permissions~\citep{Sadeh:2014vq}.
They found four clusters of different users each with different kinds of policies:
  conservatives who do not like granting many permissions,
  advanced users who seem to have more complex policies,
  unconcerned users who'll comfortably grant anything,
  and fence-sitters (about half of all users) who seemed not to really care the majority of the time.
This suggests that users are applying policies and do care how apps behave.
Further more \citeauthor{Sadeh:2014vq} speculate that the fence-sitters seemed not to care because of \emph{warning fatigue}.
This suggests a system like AppPAL that can reduce user interaction with permission warnings might be interesting.
Using a policy unconcerned and fence-sitting users could take up more privacy preserving policies without having to change their behavior.
